---
title: "Lab 5 Assignment"
author: "Hamze Omar"
date: "2021-02-18"
output: github_document
---

## Welcome to my lap 5 assignment.

- The objectives of this lab are listed below.

- Rename variables

- Create new variables

- Summarize a dataset

- Summarize a dataset by groups

- Visualize the relationship between variables

- Numerical and categorical variable

- Two numerical variables

- Three variables

Here I am going to load the packages.


```{r}
library(tidyverse)    # load the tidyverse package

library(psych)

library(scatterplot3d)
```

Here's where I read the data.

```{r}
# # read data -------------------------------------------------------------

iris
iris <- as_tibble(iris)
iris

```



## Question 1

Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.

```{r message=TRUE}
rename(iris,
  sepal_length = Sepal.Length,
  sepal_width = Sepal.Width,
  petal_lenght = Petal.Length,
  petal_width = Petal.Width
)

```

## Question 2

Convert the four numerical variables from mm to cm by multiplying by 10. Print the resulting table.

```{r message=TRUE}
mutate_each(iris, funs(. * 10), -Species)
```

## Question 3

Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.

```{r message=TRUE}
mutate(iris,
       Sepal.Area = Sepal.Width * Sepal.Length,
       Petal.Area = Petal.Width * Petal.Length)
```

## Question 4

Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table:sample size, maximum value, minimum value, range, median, first quartile (q1), third quartile (q2), inter-quartile range (iqr)


```{r message=TRUE}
summarise(
  iris,
  sample_size = n(),
  mean = (Sepal.Width),
  max = (Sepal.Width),
  median = (Sepal.Width),
  sem = mean + sqrt(sample_size)
)

scores <- c(2, 4.4, 3, 7, 4.4)
sort(scores)
min(scores)
max(scores)
median(scores)
range(scores)
quantile(scores, .25)

```

## Question 5

Calculate the following statistics for each species from the petal width variable and print the resulting table: sample size, mean, standard deviation, variance, standard error of the mean, approximate 95% confidence interval

```{r message=TRUE}
 iris_grouped <- group_by(iris, Species)
iris_grouped

iris_summarized <-
  summarise(
  iris_grouped,
  sample_size = n(),
  mean = mean(Petal.Width),
  sd = sd(Petal.Width),
  var = var(Petal.Width),
  stderr = sd / sqrt(sample_size),
  ci_upper = mean + 2 * stderr,
  ci_lower = mean - 2 * stderr
)
iris_summarized

```

## Question 6

Visualize the relationship between petal length and species using a strip plot.

```{r message=TRUE}
ggplot(data = iris) + geom_jitter(
  aes(y = Petal.Width, x = Species, color = Species),
  width = 0.05,
  height = 0.2,
  alpha = 0.5
)
```

## Question 7

Starting with the previous graph, add the mean and 95% confidence interval for each species

```{r message=TRUE}
ggplot(data = iris) + geom_jitter(
  aes(y = Petal.Width, x = Species, color = Species),
  width = 0.05,
  alpha = 0.5
) +
  geom_crossbar(
    data = iris_summarized,
    mapping = aes(
      x = Species,
      y = mean,
      ymax = ci_upper,
      ymin = ci_lower,
      color = Species
    )
  )
```

## Question 8

Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.

```{r message=TRUE}
ggplot(data = iris) + geom_jitter(
  aes(y = Petal.Width, x = Petal.Length, color = Species),
  width = 0.9,
  height = 0.5,
  alpha = 0.7
)

```


## Extra information

created question_8 as scatterplot3D, not that it was asked, but I had to try it, I guess.

```{r message=TRUE}
attach(iris)
scatterplot3d(
  Petal.Length,
  Petal.Width,
  Species,
  pch = 16,
  type = "h",
  angle = 65, box = FALSE, grid = TRUE, col.grid = "red", col.axis = "orange",
  col.lab = "navyblue",
  xlab = "Petal length",
  ylab = "Petal width",
  zlab = "Species",
  color = c("black", "purple", "blue")[as.integer(Species)]
)

```


This is the end of the lab report.

Thanks for the review.


